\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsthm,tabu,enumerate,tikz}
\usepackage[margin=1in]{geometry}
\usepackage{verbatim} % Allows Multi-line comments 
\usetikzlibrary{automata,positioning}
\newcommand{\encode}[1]{\langle #1 \rangle}

\title{Assignment 5}
\author{David Cobbley, Sasha Fahrenkopf \& Max Marchuk}
\date{December 3 2014}

\begin{document}

\maketitle


\paragraph{Problem 1} Use a reduction to show that the language $ALL_{TM}$ is undecidable
\[ALL_{TM} = \{ \langle M \rangle \;|\; Where M \mbox{ is a $TM$ and $L(M)$ = $\Sigma$*}\} \]


% I couldn't figure out a way to write the EQ proof, so I did A_TM instead
\begin{proof}
    Assume $ALL_{TM}$ is decidable. Let $R$ be a TM that decides $ALL_{TM}$.  Use $R$ to construct TM $S$ that decides $A_{TM}$.
    \begin{enumerate}
        \item $S$ = ``On input $\langle M, w \rangle$:
            \begin{enumerate}[1.]
            \item Construct $M'$ that accepts all inputs.
            \item Run $R$ on input $\langle M' \rangle$. If $R$ accepts, ACCEPT. Otherwise REJECT.''
            \end{enumerate}

        %I'm not 100% sure this is the right direction of which is reducible to which...
        Since $ALL_{TM}$ is reducible to $A_{TM}$ and $A_{TM}$ is undecidable, $ALL_{TM}$ is undecidable.
    \end{enumerate}
    \end{proof}



    
    \begin{comment}
        $S =$ ``On input $\langle M \rangle$ :
        show that $ALL_{TM}$ is undecidable
        $F =$ "on input $(\langle M \rangle , W)$:
        \begin{enumerate}
        \item Construct M'
        $M' = "On input x$:
        \begin{enumerate}[1]
        \item if $ W = \epsilon$, Reject
        \item Run M on W.
         If it accepts W, Accept X.
         If it Rejects W, Reject x"
        \end{enumerate}
        \item Output $\langle M' \rangle$"
        \end{enumerate}
        
        If M accepts W:  L(M') = $\Sigma*$
        \\If M Rejects W:  L(M') = $\emptyset$
        \\If M loops on W: L(M') = $\emptyset$
        \\--------------------------------------------------
        Other Strategy, not mapping just reduction
    
    \begin{proof}
    Assume $ALL_{TM}$ is decidable. Let $R$ be a TM that decides $ALL_{TM}$.  Use $R$ to construct TM $S$ that decides $ALL_{TM}$.
    \begin{enumerate}
        \item " On input $\langle M, M' \rangle$:
        \\ $S =$ ``On input $x$:
            \begin{enumerate}[1.]
            \item Accept X"
            \end{enumerate}
        \item Run $EQ_{TM}$ on input $\langle M \rangle and \langle M' \rangle$ 
         If $EQ_{TM}$ accepts, Accept. 
         If $EQ_{TM}$ rejects, Reject."
    \end{enumerate}
     Since $ALL_{TM}$ is reducible to $EQ_{TM}$ and $EQ_{TM}$ is undecidable, $ALL_{TM}$ is undecidable.
    \end{proof}
    \end{comment}
    
    
    

\paragraph{Problem 2} A \textit{useless state} in a Turing machine is one that is never entered on any input string. Consider the problem of determining whether a Turing Machine has any useless states. Formulate this problem as a language and show that it is undecidable."
 
%Winner
\begin{proof}  
    Assume $U_{TM}$ is decidable where $U_{TM}$ = \{ $\langle M, q \rangle$  $|$ $M$ is a TM and $q$ is a useless state\}. Let R be a TM that decides $U_{TM}$. Use $R$ to construct TM $S$ that decides $E_{TM}$.
    \begin{enumerate}
    \item S = ``On input $\langle M \rangle$:
    \begin{enumerate} [1.]
    \item Run $R$ on input $\langle M, q_{accept} \rangle$ where $q_{accept}$ is the accept state of $M$.
    \\ \\ - If $R$ accepts, ACCEPT. If $R$ rejects, REJECT.''
    \end{enumerate}
\end{enumerate}

This TM will accept only when the $L(M)$ is empty, meaning the start state is never entered on any input and is therefore a useless state. Since $E_{TM}$ is undecidable and $U_{TM}$ is reducible to $E_{TM}$, $U_{TM}$ is undecidable.   

\end{proof}
 
 
    
    %We need to mark off every state in a machine for every possible string, if any are missed its a useless 
    %State. But this is an infinite problem. Is this problem reducable somehow??

    \begin{comment}
    \begin{proof}  
    
    Assume, $U_{TM}$ is decidable. Let R = the TM that decides $U_{TM}$.
    \\S = TM that decides $E_{TM}$
    \\Proof by contradiction\\
    $HALT_{TM}$ = \{$\langle M,W \rangle$ \;|\; M halts on W \} $U_{TM} \leq_m HALT_{TM}$
    \\-Assume $U_{TM}$ is decidable. Let R = decider for $U_{TM}$
    \\-Construct S that decides $HALT_{TM} using R$.
    S=``On input $\langle M,W \rangle$:
    \begin{enumerate}
    \item Run R n $\langle M, q_{accept} \rangle$ $q_{accept} \in S$
    \\-If R accepts, S accepts
    \\-If R rejects, Reject"
    \end{enumerate}
    Because $U_{TM}$ is reducible to $HALT_{TM}$,  and $HALT_{TM}$ is undecidable, $U_{TM} is undecidable$.
    \end{proof}
    -------------------------------------------
    \end{comment}
    

    
    
\paragraph{Problem 3} If $A$ $\leq_m$ $B$ and $B$ is a regular language, does this imply that $A$
is a regular language? Why or why not?

\paragraph{} No because it is possible to create a computable function such that for every string $w$ that is an element of some context-free language $A$, the result of the function on the same input $w$ is an element of some regular language $B$. Because a context-free language can be mapping-reducible to a regular language, the difficulty class of $B$ is not guaranteed to be equal to or harder than $A$'s based on the property of regularity alone. 
\paragraph{}For example, let $A$ = \{$0^n1^n$ $|$ $n\geq 1$ \} and $B$ = \{0\}. Suppose there is some TM $M$ that is constructed with the function f: \{$0, 1\}^*$ $\rightarrow$ \{{$0, 1\}^*$ where on any input string $w$ = $0^+1^+$, the function outputs $0$. Otherwise it outputs $1$.


\paragraph{Problem 4} Prove that the language
\[LOOP_{TM} = \{ \langle M \rangle \;|\; M \mbox{ is a $TM$ and M loops on all inputs}\} \]
is not recognizable.
\begin{proof}
Assume $LOOP_{TM}$ is decidable. Let $R$ be a TM that decides for $LOOP_{TM}$.  Use $R$ to construct TM $S$ that decides $A_{TM}$.
\begin{enumerate}
\item S = ``On input $\langle M, w \rangle$:
\begin{enumerate}[1.]
\item Construct $M'$.
\\$M'$ = ``On input $x$:
\begin{enumerate}[1.]
\item Run $R$ on $w$
\\ - If $R$ accepts, ACCEPT x.
\\ - If $R$ rejects, REJECT x."
% This means that M' will only accept or reject (halt) if w never does, which is impossible!
\end{enumerate}
\end{enumerate}
\end{enumerate}
The construction of this machine proves a contradiction because $LOOP_{TM}$ is reducible to $A_{TM}$ and $A_{TM}$ is undecidable, so $LOOP_{TM}$ is also undecidable. Let $\overline{LOOP_{TM}}$ = \{ $\langle M, w \rangle$ $|$ M is a TM that loops on some inputs\}. There is a TM $T$ that recognizes $\overline{LOOP_{TM}}$ where given a string $w \in \Sigma^*$ if $w \in L(T)$, $T$ will accept and if $w \notin L(T)$ $T$ either rejects or loops. Because $LOOP_{TM}$ is undecidable and $\overline{LOOP_{TM}}$ is recognizable, $LOOP_{TM}$ is not recognizable.

% If we can prove that the complement of the language is recognizable, that will prove that the
% language is not recognizable (otherwise it would be decidable and recognizable).


\end{proof}

\end{document}
